Consumption and resistance across states and HRRs

### 2017-02-28 states/resistance/state-report.Rmd
refactoring to integrate

```{r load_data}
coherence = read_tsv('../../../../antibiogram/analysis/geographic-coherence/coherences.tsv') %>%
  rename(resistance_drug=drug) %>%
  mutate(resistance_drug=tolower(resistance_drug),
         state_coherent=level %in% c('state', 'hrr', 'zipcode'),
         hrr_coherent=level %in% c('hrr', 'zipcode')) %>%
  select(bug, resistance_drug, state_coherent, hrr_coherent)

@@

# Antibiogram data, with individual records
abg = read_tsv('../../../../antibiogram/data/abg.tsv', col_types=cols(zipcode='c')) %>%
  mutate(drug=tolower(drug)) %>%
  rename(resistance_drug=drug) %>%
  left_join(hrr, by='zipcode') %>%
  select(bug, resistance_drug, state, hrr, percent_susceptible)

@@

models = function(df, indep) {
  out = data_frame()
  for (expl in c('cpkp', 'did')) {
    new_rows = safe_spearman(df, expl, indep) %>%
      mutate(explanatory=expl)

    out = bind_rows(out, new_rows)
  }
  out
}

@@

state_res = state_dat %>%
  group_by(bug, resistance_drug, state) %>%
  summarize(mean_sus=mean(percent_susceptible, na.rm=TRUE)) %>%
  left_join(bug_drug, by=c('bug', 'resistance_drug')) %>%
  left_join(state_consumption, by=c('state', 'consumption_drug')) %>%
  mutate_at(vars(cpkp, did), function(x) if_else(is.na(x), 0, x)) %>%
  group_by(bug, resistance_drug, consumption_drug) %>%
  do(models(., 'mean_sus')) %>%
  ungroup %>%
  filter(!is.na(estimate)) %>%
  mutate(adj.p.value=p.adjust(p.value, method='BH'),
         sig=pvalue_stars(adj.p.value)) %>%
  left_join(coherence, by=c('bug', 'resistance_drug')) %>%
  mutate(record_number=1:n(),
         plot_fn=sprintf('plots/%03i.pdf', record_number),
         plot_link=pandoc.link.return(url=plot_fn, text='plot')) %>%
  select(bug, consumption_drug, resistance_drug, state_coherent,
         explanatory, correlation=estimate, p.value,
         adj.p.value, sig,
         plot_fn, plot_link)

hrr_res = hrr_dat %>%
  group_by(bug, resistance_drug, hrr) %>%
  summarize(mean_sus=mean(percent_susceptible, na.rm=TRUE)) %>%
  left_join(bug_drug, by=c('bug', 'resistance_drug')) %>%
  left_join(hrr_consumption, by=c('hrr', 'consumption_drug')) %>%
  mutate_at(vars(cpkp, did), function(x) if_else(is.na(x), 0, x)) %>%
  group_by(bug, resistance_drug, consumption_drug) %>%
  do(models(., 'mean_sus')) %>%
  ungroup %>%
  filter(!is.na(estimate)) %>%
  mutate(adj.p.value=p.adjust(p.value, method='BH'),
         sig=pvalue_stars(adj.p.value)) %>%
  left_join(coherence, by=c('bug', 'resistance_drug')) %>%
  mutate(record_number=1:n(),
         plot_fn=sprintf('plots/hrr-%03i.pdf', record_number),
         plot_link=pandoc.link.return(url=plot_fn, text='plot')) %>%
  select(bug, consumption_drug, resistance_drug, hrr_coherent,
         explanatory, correlation=estimate, p.value,
         adj.p.value, sig,
         plot_fn, plot_link)

coherent_state_res = state_res %>%
  filter(state_coherent) %>%
  mutate(adj.p.value=p.adjust(p.value, method='BH'),
         sig=pvalue_stars(adj.p.value))

coherent_hrr_res = hrr_res %>%
  filter(hrr_coherent) %>%
  mutate(adj.p.value=p.adjust(p.value, method='BH'),
         sig=pvalue_stars(adj.p.value))
```

```{r make_plots, cache=TRUE}
unlink('plots/*.pdf')

for (i in 1:nrow(state_res)) {
  row = state_res[i,]

  plot_dat = state_dat %>%
    filter(bug==row$bug, consumption_drug==row$consumption_drug)

  p = plot_dat %>%
    filter(bug==row$bug, consumption_drug==row$consumption_drug) %>%
    ggplot(aes_string(row$explanatory, 'percent_susceptible')) +
    geom_point(shape=1) +
    xlab(row$explanatory) +
    ylab('percent susceptible') +
    theme_minimal()

  ggsave(row$plot_fn, plot=p)
}

for (i in 1:nrow(hrr_res)) {
  row = hrr_res[i,]

  plot_dat = hrr_dat %>%
    filter(bug==row$bug, consumption_drug==row$consumption_drug)

  p = plot_dat %>%
    filter(bug==row$bug, consumption_drug==row$consumption_drug) %>%
    ggplot(aes_string(row$explanatory, 'percent_susceptible')) +
    geom_point(shape=1) +
    xlab(row$explanatory) +
    ylab('percent susceptible') +
    theme_minimal()

@@

First, I show the results from all correlations:

### State

```{r display_results}
state_res %>%
  mutate_if(is.numeric, function(x) signif(x, digits=3)) %>%
  select(bug, consumption_drug,
         state_coherent, explanatory, correlation,
         adj.p.value, sig,
         plot_link) %>%
  kable
```

>>>>

```{r load_data}
coherence = read_tsv('../../../../antibiogram/analysis/geographic-coherence/coherences.tsv') %>%
  rename(resistance_drug=drug) %>%
  mutate(resistance_drug=tolower(resistance_drug),
         state=level %in% c('state', 'hrr', 'zipcode'),
         hrr=level %in% c('hrr', 'zipcode')) %>%
  select(bug, resistance_drug, state, hrr) %>%
  gather('place', 'coherent', state, hrr)

@@

# Antibiogram data, with individual records
abg = read_tsv('../../../../antibiogram/data/abg.tsv', col_types=cols(zipcode='c')) %>%
  mutate(drug=tolower(drug), percent_nonsusceptible=100-percent_susceptible) %>%
  rename(resistance_drug=drug) %>%
  left_join(hrr, by='zipcode') %>%
  select(bug, resistance_drug, state, hrr, percent_nonsusceptible)

@@

models = function(df, indep, keep=NULL) {
  lapply(c('cpkp', 'did'), function(expl) {
    safe_spearman(df, expl, indep) %>%
      mutate(explanatory=expl)
  }) %>%
    bind_rows
}

@@

models_result = function(x, this_place, consumption) {
  result = x %>%
    group_by_('bug', 'resistance_drug', this_place) %>%
    summarize(mean_nonsus=mean(percent_nonsusceptible, na.rm=TRUE)) %>%
    left_join(bug_drug, by=c('bug', 'resistance_drug')) %>%
    left_join(consumption, by=c(this_place, 'consumption_drug')) %>%
    mutate_at(vars(cpkp, did), function(x) if_else(is.na(x), 0, x)) %>%
    group_by(bug, resistance_drug, consumption_drug) %>%
    do(models(., 'mean_nonsus')) %>%
    ungroup %>%
    filter(!is.na(estimate)) %>%
    mutate(adj.p.value=p.adjust(p.value, method='BH'),
           sig=pvalue_stars(adj.p.value)) %>%
    left_join(filter(coherence, place==this_place), by=c('bug', 'resistance_drug')) %>%
    mutate(record_number=1:n(),
           plot_fn=sprintf('plots/%s_%03i.pdf', this_place, record_number),
           plot_link=pandoc.link.return(url=plot_fn, text='plot')) %>%
    select(bug, consumption_drug, resistance_drug, coherent,
           explanatory, correlation=estimate, p.value,
           adj.p.value, sig,
           plot_fn, plot_link)

  # add adjusted p-values for just the coherent stuff
  # swo> there's probably a more elegant way to do this
  coherent.adj.p.value = result %>%
    filter(coherent) %$%
    p.adjust(p.value, method='BH')
  coherent.sig = pvalue_stars(coherent.adj.p.value)
  result$coherent.adj.p.value = NA
  result$coherent.sig = NA
  result$coherent.adj.p.value[result$coherent] = coherent.adj.p.value
  result$coherent.sig[result$coherent] = coherent.sig

  result
}

state_res = models_result(state_dat, 'state', state_consumption)
hrr_res = models_result(hrr_dat, 'hrr', hrr_consumption)

result = bind_rows(
  state_res %>% mutate(place='state'),
  hrr_res %>% mutate(place='hrr')
)
```

```{r make_plots}
unlink('plots/*.pdf')

for (i in 1:nrow(result)) {
  row = result[i,]

  if (row$place == 'state') {
    plot_dat = state_dat
  } else if (row$place == 'hrr') {
    plot_dat = hrr_dat
  } else {
    plot_dat = NULL
  }

  p = plot_dat %>%
    filter(bug==row$bug, consumption_drug==row$consumption_drug) %>%
    ggplot(aes_string(row$explanatory, 'percent_nonsusceptible')) +
    geom_point(shape=1) +
    xlab(row$explanatory) +
    ylab('percent non-susceptible') +
    theme_minimal()

@@

Here a relationship where increased consumption leads to increased nonsusceptibility is encoded in a positive correlation.

### Geography-averaged results

In these results, the percent nonsusceptible in each geography (state or HRR) was averaged, and the averages of the nonsusceptibilities were compared against the consumption data.

#### State-level geography-averaged correlations

```{r state_average_results_table}
results %>%
  filter(place=='state') %>%
  mutate_if(is.numeric, function(x) signif(x, digits=3)) %>%
  select(bug, consumption_drug,
         coherent, explanatory, correlation,
         adj.p.value, sig,
         plot_link) %>%
  kable
```

There correlations are mostly positive, but only a small number are significant (red line marks $p = 0.05$):

```{r state_average_histogram}
result %>%
  filter(place=='state') %>%
  ggplot(aes(x=correlation, y=adj.p.value)) +
  geom_point() +
  geom_hline(yintercept=0.05, col='red') +
  xlab('correlation coefficient') + ylab('p value (reversed)') +
  scale_y_reverse() +
  theme_minimal()
```
